---
title: "QualityControlForMaxQuantOutput"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{QualityControlForMaxQuantOutput}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = F}
library(QCCFR)
```

# Performing QC of proteinGroups.txt or Phospho (STY)sites.txt
## 1. Import table.

There are two functions, one for each proteinGroups and STY
```{r eval = F}
proteinGroups <- 
  read_proteinGroups(
    folderPath = "~/data",    # Set this to wherever your data is saved
    
    pattern = "^P_",           # In the example data, the proteome samples
    # all have "P_" at the start of the column name
    # Often our experiments will have "STY" or "PRO" in the
    # name that can ber used here to differentiate.
    quantification = "LFQ"   # This was a label free experiment
  )
# If this was a SILAC experiment, I would also optionally set the parameter
# SILAC ratios, to specify which columns I'm interested in. See the
# documentation for details.

phosphoSTYsites <- 
  read_phosphoSTY(
    folderPath = "~/data",    # Set this to wherever your data is saved
    
    pattern = "^LPS",# In the example data, the phosphoproteome samples
    # all have "LPS" at the start of the column name.
    # Often our experiments will have "STY" or "PRO" in the
    # name that can ber used here to differentiate.
    quantification = "LFQ"  
  )

```

## 2. Filter the table
Proteins/sites are removed if they match the following criteria\
- **Potential Contaminants** \
- **Reverse** i.e. peptides that match the decoy peptides in the search engine\
- **Proteins/sites identified but not quantified** (incomplete cases)\
Protein Groups only:\
- **Only identified by site** (for protein groups)\
- **Razor / unique peptides >1**\
- **`Unique + razor sequence coverage [%]` >= 5**\
```{r}
proteinGroups_flt <- 
  filter_proteinGroups(
    proteinGroups,            # Output of read_proteinGroups()
    quantification = "LFQ"
  )
```
Phospho(STY) only:\
- **Localisation probability > 0.7**\
```{r}
phosphoSTYSites_flt <- 
  filter_sty(
    phosphoSTYSites,            # Output of read_phosphoSTY()
    quantification = "LFQ"
  )
```

## 3. Normalise the data
All the following functions work the same regardless of whether you pass a proteinGroups or STY table to them :)
```{r}
proteinGroups_flt_norm <- 
  normaliseData(
    proteinGroups_flt,      # Output of filter_proteinGroups
    quantification = "LFQ"
  )
```

`proteinGroups_flt_norm` is a list with two components:\

* **dat** a dataframe of the normalised data\
* **plots** two plots, one eac of the pre-normalised and post-normalised data
```{r echo = F, results = "asis"}
knitr::kable(head(proteinGroups_flt_norm$dat))
proteinGroups_flt_norm$plots
```

### Steps 1-3 can also be done as a dplyr pipe:
```{r eval = FALSE}
read_proteinGroups(
  folderPath = "~/data",   
  pattern = "^P_",        
  quantification = "LFQ"  
) %>% 
  filter_proteinGroups(quantification = "LFQ") %>% 
  normaliseData(quantification = "LFQ")
```

## 4. Visualisation
```{r}
pca_plot(proteinGroups_flt_norm,
         quantification = "LFQ",
         colour_grp_pattern = "_0[0123]"  # Using regex, group your 
         # experimental conditions.
         # Along with shape_grp_pattern and colour_scale, these are
         # optional parameters!
)
```

```{r}
cor_plot(proteinGroups_flt_norm,
         quantification = "LFQ")
```
```{r}
count_plot(proteinGroups_flt_norm,
           quantification = "LFQ")
```

# Performing QC ofmodificationSpecificPeptides.txt
## 1. Read data
```{reval = F}
modsp <- read_modSpPeptides(path, 
                            quantification = "LFQ")
```

## 2. Visualise characteristics of phosphorylated peptides
```{r}
pies <- modificationCharacteristics(
  phosphoSTYSites,                      # Note that this takes STY table also
  modificationSpecificPeptides, 
  quantification = "LFQ"
  )
pies
```

Note: You can use the `patchwork` library to join these plots into one graphic:
```{r eval = F}
library(patchwork)
pies[[1]] + pies[[2]] + pies[[3]]
```

## 3. Calculate enrichment efficiency of phosphorylated peptides:
```{r}
phosphoenrichment(modificationSpecificPeptides,
                  quantification = "LFQ",
                  pattern = " LPS_")
```



# Evidence.txt
```{r eval = F}
evidence <- read_evidence(
  folderPath = "~/data"
)
```

```{r}
massErrorDensity(evidence) # This looks more impressive with more than 50
# peptides to visualise :)
andromedaScorePlot(evidence)
```

If you just want a subset of the samples you can specify the required experiments using **pattern** which takes a regex string argument:
```{r}
massErrorDensity(evidence, pattern = " LPS_")
```
